Interval Probability and Measurement for Fuzzy Quantum Theories

### 2016-08-18 prob.tex
prob.tex unifying the definitions...

The conventional definition of a probability space builds upon the
field of real numbers. In more detail, a probability space consists
of a \emph{sample space} $\Omega$, a space of \emph{events}~$\events$,
and a \emph{probability measure}~$\pmeas$ mapping events in $\events$
to the real interval $[0,1]$. In this paper, we will only consider
\emph{finite} sets of events: we therefore restrict our attention
to non-empty finite sets $\Omega$ as the sample space. The space
of events $\events$ includes every possible subset of $\Omega$:
it is the powerset~$2^{\Omega}=\left\{ E\middle|E\subseteq\Omega\right\} $.
Given the set of events $\events$, a \emph{probability measure} is
a function $\pmeas:\events\rightarrow[0,1]$ such that: 
\begin{itemize}
\item $\pmeas(\Omega)=1$, and 
\item for a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, $\pmeas(\bigcup_{i=1}^{N}E_{i})=\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\pmeas(E_{i})$,
where $\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\pmeas(E_{i})$ explicitly
specifies $\pmeas(E_{i})\in\mathbb{R}$. Besides $\mathbb{R}$, we
will prepose other symbols later to specify the type of operations,
and they may be dropped when there is no ambiguity. 
\end{itemize}

@@

\begin{lemma}\label{lem:classicalProbabilityMeasure}Given a sample
space~$\Omega$ and its event~$\events$, a function~$\mu:\events\rightarrow[0,1]$
is a classical interval-valued probability measure if and only if
$\mu$ satisfies the following conditions: 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$, 
\item $\bar{\mu}(\Omega)=[1,1]$, 
\item for any event $E$, $\bar{\mu}\left(\Omega\backslash E\right)=\left[1,1\right]\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(E\right)$,
where $\Omega\backslash E$ is the complement event, and 
\item for a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have $\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)\subseteq\tensor*[_{\mathscr{I}}]{\sum}{_{i=1}^{N}}\bar{\mu}\left(E_{i}\right)$,
where $\tensor*[_{\mathscr{I}}]{\sum}{_{i=1}^{N}}[l_{i},r_{i}]=\left[\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}l_{i},\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}r_{i}\right]$.
We may drop the preposing $\mathscr{I}$ when summands are clearly
intervals. 
\end{itemize}
\qed\end{lemma}

@@

We present an example before giving the formal definition of quantum
probability measures.

@@

%%%%%
\subsection{Quantum Real-valued Probability Measures}

\begin{definition}[Quantum Probability Measure~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}]\label{def:QuantumProbabilitySpace}
Given a Hilbert space $\Hilb$ with its set of events~$\events$,
a \emph{quantum probability measure} is a function~$\mu:\events\rightarrow[0,1]$
such that:\footnote{It is possible to define a more general space of events consisting
of all operators~$\mathcal{A}$ on $\Hilb$ and to extend $\mu$
linearly to $\mathcal{A}$ and consider $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$
is the expectation value of $A$ which could bigger than one because
$\mu$ is linear. We does not take this approach because we want to
focus only on probability. } 
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\mu\left(\tensor*[_{\mathscr{O}}]{\sum}{_{i=1}^{N}}P_{i}\right)=\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\mu\left(P_{i}\right)$. 
\end{itemize}
\qed\end{definition}

In order to motivate the definition of quantum interval-valued probability
measures later, we provide an equivalent definition of quantum probability
measures.

\begin{lemma}\label{lem:quantumProbabilityMeasure}Given a Hilbert
space $\Hilb$, a function~$\mu:\events\rightarrow[0,1]$ is a quantum
probability measure if and only if $\mu$ satisfies the following
conditions: 
\begin{itemize}
\item $\mu(\mathbb{0})=0$, 
\item $\mu(\mathbb{1})=1$, 
\item for any projection $P$, $\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)$,
and 
\item for a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\mu\left(\tensor*[_{\mathscr{O}}]{\sum}{_{i=1}^{N}}P_{i}\right)=\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\mu\left(P_{i}\right)$. 
\end{itemize}
\qed\end{lemma}

A set of events~$\events$ together with a quantum probability measure
is called a \emph{quantum probability space}. Comparing to classical
probability spaces, the empty set~$\emptyset$ corresponds to the
empty projection~$\mathbb{0}$, and the event of whole space~$\Omega$
corresponds to the identity projection~$\mathbb{1}$. In contrast,
the union~$\cup$ of any two events always gives an event classically,
but the operator addition~$\tensor*[_{\mathscr{O}}]{+}{}$ of two
projections may not be a projection. As the result, the classical
condition~$\pmeas(E_{1}\cup E_{2})=\pmeas(E_{1})\tensor*[_{\mathbb{R}}]{+}{}\pmeas(E_{2})$
is always defined, and it is true when $E_{1}$ and $E_{2}$ are disjoint;
however, $\mu\left(P_{1}\tensor*[_{\mathscr{O}}]{+}{}P_{2}\right)=\mu(P_{1})\tensor*[_{\mathbb{R}}]{+}{}\mu(P_{2})$
is always true whenever the left-handed side is defined.

@@

Notice that the mixed state~$\rho$ is really a generalization of
Radon-Nikodym derivative if we pick a basis~$\left\{ \ket{\psi_{j}}\right\} _{j=1}^{d}$
such that $P=\tensor*[_{\mathscr{O}}]{\sum}{_{\ket{\psi_{j}}\in S}}\proj{\psi_{j}}$,
where $S$ is the linear subspace $P$ projects onto. By representing
the density matrix~$\rho$ in the same basis, i.e., $\rho=\sum_{i=1}^{d}\sum_{j=1}^{d}r_{ij}\ket{\psi_{i}}\bra{\psi_{j}}$,
the Born rule becomes
\begin{eqnarray}
\mu_{\rho}^{B}\left(P\right) & = & \Tr\left(\rho P\right)=\sum_{\ket{\psi_{j}}\in S}r_{jj}\textrm{ ,}\label{BornRule.mixed-1}
\end{eqnarray}
 which is basically the same as equation (\ref{eq:Radon-Nikodym}).

In this sense, the Radon-Nikodym theorem could be generalized to Gleason's
theorem~\cite{HollandJr1970,Redhead1987-REDINA,Jaeger2007}: in Hilbert
spaces of dimension $d\geq3$, given a quantum probability measure
$\mu$, there exist a mixed state~$\rho$ that induces such a measure
using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
By applying Gleason's theorem, we can extend lemma~\ref{lem:quantumProbabilityMeasure}
to the following corollary.

\begin{cor}\label{cor:Gleason's}Given a Hilbert space $\Hilb$ of
dimension $d\geq3$, for any function~$\mu:\events\rightarrow[0,1]$
satisfying the conditions listed in lemma~\ref{lem:quantumProbabilityMeasure},
there exists an unique mixed state~$\rho$ such that $\mu=\mu_{\rho}^{B}$.\end{cor}

@@

As mentioned above, each quantum event~$P$ will be associated with
an interval-valued probability~$\left[l,r\right]$. Notice that whenever
the magnetic field of the Stern-Gerlach is fixed, i.e., an ideal measurement
is picked, measuring the spin is exactly the same as tossing a coin.
In another word, if somebody claims she is tossing a coin behind a
veil, and only shows the resulting heads or tails, we cannot distinguish
whether she has really tossed a coin, or she has run an Stern-Gerlach
experiment, and shows us the head, the tail, or the side of coin if
the silver atoms is spin up, spin down, or just hit the middle, respectively.
Therefore, we should plug in the definition of interval-valued probability
measure into the definition of quantum probability space for each
ideal measurement, and get the following definition.

>>>>

The conventional definition of a probability space builds upon the
field of real numbers. In more detail, a probability space consists
of a \emph{sample space} $\Omega$, a space of \emph{events}~$\events$,
and a \emph{probability measure}~$\pmeas$ mapping events in $\events$
to the real interval $[0,1]$. In this paper, we will only consider
\emph{finite} sets of events: we therefore restrict our attention
to non-empty finite sets $\Omega$ as the sample space. The space
of events $\events$ includes every possible subset of $\Omega$:
it is the powerset~$2^{\Omega}=\left\{ E\middle|E\subseteq\Omega\right\} $.

\begin{definition}[Probability Measure]\label{def:ClassicalProbabilitySpace}
Given the set of events $\events$, a \emph{probability measure} is
a function $\pmeas:\events\rightarrow[0,1]$ such that: 
\begin{itemize}
\item $\pmeas(\emptyset)=0$. 
\item $\pmeas(\Omega)=1$. 
\item for any event $E$, 
\begin{equation}
\pmeas\left(\Omega\backslash E\right)=1\tensor*[_{\mathbb{R}}]{-}{}\pmeas\left(E\right)\textrm{ ,}\label{eq:ClassicalReal-valuedProbability-Complement}
\end{equation}
where $\Omega\backslash E$ is the complement event, and $1\tensor*[_{\mathbb{R}}]{-}{}\pmeas\left(E\right)$
explicitly specifies $1$ and $\pmeas\left(E\right)\in\mathbb{R}$.
Besides $\mathbb{R}$, we will prepose other symbols later to specify
the type of operations, and they may be dropped when there is no ambiguity. 
\item for a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, $\pmeas(\bigcup_{i=1}^{N}E_{i})=\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\pmeas(E_{i})$.
Similar to the above condition, $\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\pmeas(E_{i})$
explicitly specifies $\pmeas(E_{i})\in\mathbb{R}$. 
\end{itemize}
\qed\end{definition}

Notice that $\pmeas(\emptyset)=0$ and equation~(\ref{eq:ClassicalReal-valuedProbability-Complement})
can be induced by other conditions. They are listed in here so that
we can compare definition~\ref{def:ClassicalProbabilitySpace}, lemma~\ref{lem:classicalProbabilityMeasure},
definition~\ref{def:QuantumProbabilitySpace}, and definition~\ref{def:QuantumInterval-valuedProbability}
easily. We will compare these lemma and definitions when we try to
formulate quantum interval-valued probability measures later.

@@

\begin{lemma}\label{lem:classicalProbabilityMeasure}Given a sample
space~$\Omega$ and its event~$\events$, a function~$\mu:\events\rightarrow[0,1]$
is a classical interval-valued probability measure if and only if
$\mu$ satisfies the following conditions: 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$, 
\item $\bar{\mu}(\Omega)=[1,1]$, 
\item for any event $E$, $\bar{\mu}\left(\Omega\backslash E\right)=\left[1,1\right]\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(E\right)$,
and 
\item for a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have $\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)\subseteq\tensor*[_{\mathscr{I}}]{\sum}{_{i=1}^{N}}\bar{\mu}\left(E_{i}\right)$,
where $\tensor*[_{\mathscr{I}}]{\sum}{_{i=1}^{N}}[l_{i},r_{i}]=\left[\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}l_{i},\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}r_{i}\right]$.
We may drop the preposing $\mathscr{I}$ when summands are clearly
intervals. 
\end{itemize}
\qed\end{lemma}

@@

@@

Based on the above definition and properties of quantum events, we
can present an example before giving the formal definition of quantum
probability measures.

\subsection{Quantum Real-valued Probability Measures}

\begin{definition}[Quantum Probability Measure~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}]\label{def:QuantumProbabilitySpace}
Given a Hilbert space $\Hilb$ with its set of events~$\events$,
a \emph{quantum probability measure} is a function~$\mu:\events\rightarrow[0,1]$
such that:\footnote{It is possible to define a more general space of events consisting
of all operators~$\mathcal{A}$ on $\Hilb$ and to extend $\mu$
linearly to $\mathcal{A}$ and consider $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$
is the expectation value of $A$ which could bigger than one because
$\mu$ is linear. We does not take this approach because we want to
focus only on probability. } 
\begin{itemize}
\item $\mu(\mathbb{0})=0$, 
\item $\mu(\mathbb{1})=1$, 
\item for any projection $P$, 
\begin{equation}
\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)\textrm{ ,}\label{eq:QuantumReal-valuedProbability-Complement}
\end{equation}
and 
\item for a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\mu\left(\tensor*[_{\mathscr{O}}]{\sum}{_{i=1}^{N}}P_{i}\right)=\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\mu\left(P_{i}\right)$. 
\end{itemize}
\qed\end{definition}

Similar to classical definition~\ref{def:ClassicalProbabilitySpace},
$\mu(\mathbb{0})=0$ and equation~(\ref{eq:QuantumReal-valuedProbability-Complement})
can be induced by other conditions. A set of events~$\events$ together
with a quantum probability measure is called a \emph{quantum probability
space}. Comparing to classical probability spaces, the empty set~$\emptyset$
corresponds to the empty projection~$\mathbb{0}$, and the event
of whole space~$\Omega$ corresponds to the identity projection~$\mathbb{1}$.
In contrast, the union~$\cup$ of any two events always gives an
event classically, but the operator addition~$\tensor*[_{\mathscr{O}}]{+}{}$
of two projections may not be a projection. As the result, the classical
condition~$\pmeas(E_{1}\cup E_{2})=\pmeas(E_{1})\tensor*[_{\mathbb{R}}]{+}{}\pmeas(E_{2})$
is always defined, and it is true when $E_{1}$ and $E_{2}$ are disjoint;
however, $\mu\left(P_{1}\tensor*[_{\mathscr{O}}]{+}{}P_{2}\right)=\mu(P_{1})\tensor*[_{\mathbb{R}}]{+}{}\mu(P_{2})$
is always true whenever the left-handed side is defined.

@@

Notice that the mixed state~$\rho$ is really a generalization of
Radon-Nikodym derivative if we pick a basis~$\left\{ \ket{\psi_{j}}\right\} _{j=1}^{d}$
such that $P=\tensor*[_{\mathscr{O}}]{\sum}{_{\ket{\psi_{j}}\in S}}\proj{\psi_{j}}$,
where $S$ is the linear subspace $P$ projects onto. By representing
the density matrix~$\rho$ in the same basis, i.e., $\rho=\sum_{i=1}^{d}\sum_{j=1}^{d}r_{ij}\ket{\psi_{i}}\bra{\psi_{j}}$,
the Born rule becomes
\begin{eqnarray}
\mu_{\rho}^{B}\left(P\right) & = & \Tr\left(\rho P\right)=\sum_{\ket{\psi_{j}}\in S}r_{jj}\textrm{ ,}\label{BornRule.mixed-1}
\end{eqnarray}
which is basically the same as equation~(\ref{eq:Radon-Nikodym}).

In this sense, the Radon-Nikodym theorem could be generalized to Gleason's
theorem~\cite{HollandJr1970,Redhead1987-REDINA,Jaeger2007}: in Hilbert
spaces of dimension $d\geq3$, given a quantum probability measure
$\mu$, there exist a mixed state~$\rho$ that induces such a measure
using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
By applying Gleason's theorem, we can extend definition~\ref{def:QuantumProbabilitySpace}
to the following corollary.

\begin{cor}\label{cor:Gleason's}Given a Hilbert space $\Hilb$ of
dimension $d\geq3$, for any function~$\mu:\events\rightarrow[0,1]$
satisfying the conditions listed in definition~\ref{def:QuantumProbabilitySpace},
there exists an unique mixed state~$\rho$ such that $\mu=\mu_{\rho}^{B}$.\end{cor}

@@

As mentioned above, each quantum event~$P$ will be associated with
an interval-valued probability~$\left[l,r\right]$. Notice that whenever
the magnetic field of the Stern-Gerlach is fixed, i.e., an ideal measurement
is picked, measuring the spin is exactly the same as tossing a coin.
In another word, if somebody claims she is tossing a coin behind a
veil, and only shows the resulting heads or tails, we cannot distinguish
whether she has really tossed a coin, or she has run an Stern-Gerlach
experiment, and shows us the head, the tail, or the side of coin if
the silver atoms is spin up, spin down, or just hit the middle, respectively.
Therefore, we should plug in the definition of interval-valued probability
measure into the definition of quantum probability space for each
ideal measurement.

More specifically, for classical probability measures, definition~\ref{def:ClassicalProbabilitySpace}
and lemma~\ref{lem:classicalProbabilityMeasure} are almost the same
except the last condition using the equal sign~$=$ in real-valued
definition~\ref{def:QuantumProbabilitySpace} and the subset equal
sign~$\subseteq$ in interval-valued lemma~\ref{lem:classicalProbabilityMeasure}.
Therefore, for quantum probability measures, replacing the the equal
sign~$=$ in real-valued definition~\ref{def:QuantumProbabilitySpace}
by the subset equal sign~$\subseteq$ gives the only reasonable definition
for quantum interval-valued probability measures as follow.
